Reproducible Research Project 1
=========================================================================================================


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First read the file and insert the data into a dataframe called stepdata:

```{r}
stepdata <- read.csv("activity.csv", sep = ",", header = TRUE)
```

## What is the mean total number of steps taken per day? 

The mean total number of steps taken per day:

```{r}
sumsteps <- aggregate(. ~ date, data = stepdata, FUN = sum)
hist(sumsteps$steps, xlab = "Steps", main = "Histogram of Steps")
meansteps <- mean(sumsteps$steps)
mediansteps <- median(sumsteps$steps)
```

The mean steps per day is `r meansteps` and the median steps is `r mediansteps`.

## What is the average daily activity pattern?

```{r}
stepinterval <- aggregate(steps ~ interval, stepdata, mean)
plot(stepinterval$interval, stepinterval$steps, type = "l", xlab = "Interval", ylab = "Steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
maxinterval <- stepinterval[which.max(stepinterval$steps),1]
```

The `r maxinterval` interval contains the maximum number of steps.

## Imputing Missing Values

Calculate the number of NA values in the data set:
```{r}
numberofNA <- sum(is.na(stepdata))
```
There are `r numberofNA` NA values in the data set.

Impute missing data by average steps of each interval:
```{r}
imputedata <- transform(stepdata, steps = ifelse(is.na(stepdata$steps), stepinterval$steps[match(stepdata$interval, stepinterval$interval)], stepdata$steps))
newsumsteps <- aggregate(steps ~ date, imputedata, sum)
```

New histogram of total number of steps taken each day.
```{r}
hist(newsumsteps$steps, main = "Total Steps Each Day", xlab = "Number of Steps", col = "blue")
```

New mean and new median for imputed data:
```{r}
newmean <- mean(newsumsteps$steps)
newmedian <- median(newsumsteps$steps)
```

The new mean is `r newmean` and the new median is `r newmedian`.  The old and new values are very close to each other.  So imputing data had a minimal effect on the previous data.


